% Abstract KBS v3 (2026-07-18): faithful compression of the IEEE abstract
% (447 words) under the Elsevier/KBS hard limit of 250 words, WITH the
% contextual opening (NER = foundational NLP task) requested by Thibault.
% Every claim of the IEEE version is kept (trade-off positioning,
% transfer-learning design, gold vs end-to-end separation, order-of-magnitude
% efficiency gaps, both variants leading their category, detection bottleneck).
% Same headline numbers: means over 3 retraining seeds (7/42/123): 62.3 gold
% (sup.), 39.5 e2e (ZS fusion), 40.1 e2e (sup.), 1.7 Wh (seed-42 run).
% No references, no undefined abbreviations (KBS abstract rules).
Named Entity Recognition (NER), a foundational Natural Language Processing task, extracts entities (persons, dates, locations) from raw text. Yet, real-world applications rarely afford a static class inventory.
Open-world NER therefore assigns types not fixed at training time, by training a dedicated encoder or prompting billion-parameter language models, compared on accuracy alone, ignoring energy cost.
We present OPENER, an open-world NER pipeline redefining the optimal trade-off between accuracy, latency, and energy frugality.
Uncommonly, this transfer-learning pipeline trains no encoder from scratch: a frozen mention detector, a Matryoshka-truncatable embedder sharpened by contrastive fine-tuning with error-driven hard-negative mining, and a swappable typing head.
OPENER-ZS, the zero-shot head, types mentions against label-name prototypes without target labels, while OPENER-Sup fits a lightweight supervised probe on labelled classes, trading openness for accuracy.
We benchmark OPENER against GLiNER, GNER, OWNER and a 4-bit Qwen2.5-1.5B model on 13 datasets along three axes (Adjusted Mutual Information (AMI), latency, energy and CO\textsubscript{2}), separating gold-mention typing from end-to-end, exposing order-of-magnitude gaps that accuracy-only evaluations hide.
Over three retraining seeds, both variants lead their category: OPENER-ZS is the most accurate zero-shot system end-to-end ($39.5$ AMI vs $38.9$ for GLiNER-L, $34.3$ for OWNER) while staying frugal ($1.7$\,Wh, $180$\,ms, versus $14.3$\,Wh, $616$\,ms for OWNER).
With target labels, OPENER-Sup is the most accurate overall ($40.1$ AMI).
On gold mentions, typing alone reaches $62.3$ AMI versus OWNER's $43.0$, pinning the open-world bottleneck on mention detection, the clearest axis for future gains.
We release code, configurations, fine-tuned embedders, and benchmark.
