% =====================================================================
% Elsevier-required closing sections (KBS Guide for Authors).
% Order: CRediT -> competing interest -> funding -> generative AI
% declaration -> data availability, all placed before the reference list.
% Note: at submission, the competing-interest declaration must ALSO be
% completed in Elsevier's online declarations tool (separate .docx upload).
% =====================================================================

\section*{CRediT authorship contribution statement}
\label{sec:credit}
\textbf{Thibault Garel:} Conceptualization, Methodology, Software, Validation, Formal analysis, Investigation, Data curation, Writing -- original draft, Writing -- review \& editing, Visualization.
\textbf{Faiza Belbachir:} Supervision.
\textbf{Assia Soukane:} Supervision.

\section*{Declaration of competing interest}
\label{sec:competing}
The authors declare no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.

\section*{Funding}
\label{sec:funding}
This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.

\section*{Declaration of generative AI and AI-assisted technologies in the manuscript preparation process}
\label{sec:genai}
During the preparation of this work the first author used Claude Code (Anthropic) in order to scaffold and debug the experiment scripts, to draft and copy-edit the manuscript, and to assist with exploratory data analysis.
The research questions, the experimental design, the choice of baselines and datasets, the execution of all experiments, and the interpretation of the results are the first author's own.
After using this tool, the authors reviewed and edited the content as needed and take full responsibility for the content of the published article.

\section*{Data availability}
\label{sec:data}
All datasets used in this study are publicly available benchmarks, cited in \autoref{sec:exp:datasets}.
The code and the per-experiment configurations are openly available on GitHub (\url{https://github.com/Thibault-GAREL/LyRIDS_OPENER}), and the fine-tuned embedders of OPENER-ZS and OPENER-Sup are available on Hugging Face (\url{https://huggingface.co/Thibault-GAREL/opener-zs} and \url{https://huggingface.co/Thibault-GAREL/opener-sup}).
