% Abstract v4 (2026-07-08, multi-seed). Energy-aware framing, headline = OPENER-ZS
% (fusion) as the best zero-shot system end-to-end. supervised probe = upper bound.
% Numbers = means over 3 retraining seeds (7/42/123): 62.3 gold (sup.), 39.5 e2e
% (ZS fusion), 40.1 e2e (sup.), 1.7 Wh (seed-42 run).
Named Entity Recognition (NER) is a foundational task in Natural Language Processing, used to extract entities from raw text. However, real-world NER applications rarely afford the luxury of a static, predefined set of entity classes. This limitation has driven the emergence of open-world NER, where systems must assign entity types that were not fixed at training time. To tackle this challenge, recent systems either train a dedicated encoder or prompt billion-parameter language models, and they are compared almost exclusively on accuracy, ignoring their compute and energy cost.
We present OPENER, an open-world NER pipeline that redefines the optimal trade-off between accuracy, latency, and energy frugality. OPENER is assembled from ready-to-use components: a frozen mention detector, a Matryoshka-truncatable sentence embedder sharpened by a light contrastive fine-tuning with error-driven hard-negative mining, and a swappable typing head, so that no encoder is trained from scratch.
This design relies on transfer learning: it reuses pretrained components instead of training a dedicated encoder or prompting a large language model, a choice still uncommon in open-world NER.
The typing head comes in two forms. OPENER-ZS is \emph{zero-shot}: it types each mention against label-name prototypes and needs no target labels, for the truly open-world setting where entity types are not known in advance. OPENER-Sup instead fits a lightweight supervised probe when a domain's entity classes are fixed and labelled, trading openness for accuracy.
We benchmark OPENER against GLiNER, GNER, OWNER and a 4-bit Qwen2.5-1.5B large language model on 13 datasets along three axes at once: quality (measured by Adjusted Mutual Information, AMI), inference latency (median per-sentence p50), and energy consumption (watt-hours and grams of CO\textsubscript{2}eq per run).
Throughout, we separate entity typing on gold mentions (given perfect mention boundaries) from the full end-to-end setting to analyze each component. Moreover, our energy-aware comparison exposes order-of-magnitude efficiency gaps that a quality-only evaluation would hide.
On our 13-dataset benchmark, both variants achieve the highest scores in their category, averaged over three embedder-retraining seeds. OPENER-ZS is the most accurate \emph{zero-shot} system end-to-end ($39.5$ AMI, all AMI scores being reported $\times 100$ as in the tables, ahead of GLiNER-L's $38.9$ and OWNER's $34.3$) while staying frugal, at $1.7$\,Wh and $\sim$$180$\,ms per sentence: several times less energy and latency than OWNER ($14.3$\,Wh, $616$\,ms), and far below the 4-bit Qwen LLM ($11.4$\,Wh, $5$\,s). Given a domain's labels, OPENER-Sup is the most accurate system overall ($40.1$ AMI*100). On gold mentions, with detection removed, typing alone reaches $62.3$ AMI versus OWNER's $43.0$: the wide gap to the end-to-end scores pins the open-world bottleneck on mention detection rather than typing, the clearest axis for future gains.
% Camera-ready: the release links (GitHub + HuggingFace) live in sections/03_method.tex,
% end of the Evaluation Protocol subsection. The real URLs are kept in comments there.
We release the code, the per-experiment configurations, the fine-tuned embedders, and the full benchmark.
