\section{Introduction}
\label{sec:introduction}

Named Entity Recognition (NER) is a critical and challenging task in the field of Natural Language
Processing (NLP). Its primary goal is to locate and classify key elements in unstructured text into predefined categories (e.g., extracting "Marie Curie" as a PERSON or "1898" as a DATE). Its most widespread uses include automating the analysis of legal or medical documents, intelligently sorting customer support tickets, or even searching and finding this research paper itself. However, real-world NER rarely operates with a static label set, as novel categories inevitably emerge across new domains. Although traditional models accurately assign contiguous sequences of tokens, called spans, into fixed classes, this rigid training renders them blind to unobserved types, like identifying "radium" as a CHEMICAL \cite{nguyen2023span}.
To avoid the slow and costly process of re-annotating data and retraining from scratch, \emph{open-world} NER is designed to dynamically recognize and type entities without relying on a frozen training inventory.

Here is some famous example of open-world detectors: GLiNER \cite{zaratiana2024gliner} scores candidate spans against the target type names supplied as queries in a single forward pass, while GNER \cite{ding2024gner} recasts NER as text generation with a sequence-to-sequence model. Both can type entities whose categories were not fixed at training time, at the cost of training a dedicated model for the task.
A complementary line prompts or instruction-tunes billion-parameter language models \cite{zhou2024universalner, sainz2024gollie} are accurate but slow, brittle, and often too large for commodity hardware. Even a small quantized instruction-tuned model that does fit, such as Qwen2.5-1.5B-Instruct \cite{qwen2025} in 4-bit which we evaluate, stays comparatively slow and costly.
OWNER \cite{genest2025owner} takes yet another route, training a dedicated entity encoder with a triplet objective and clustering the resulting mentions into dynamically named types. But it still trains an encoder from scratch and pays a heavy per-sentence inference cost.

Across all of these, evaluation is reported almost exclusively on accuracy: the compute, latency and energy consumption that separate a $467$\,M-parameter pipeline from a $7$\,B-parameter model are left out of the comparison, even though they decide what can actually be deployed.

We present \textbf{OPENER}, an open-world NER pipeline built for all three axes at once: on our $13$-dataset benchmark, it is the \textbf{most accurate} of the compared systems while remaining \textbf{fast} and \textbf{energy-frugal}. Unusually, it trains no encoder from scratch. It is assembled entirely from ready-to-use and pretrained components, a \textbf{design still uncommon} in open-world NER.
Concretely, a frozen \textbf{mention detector} proposes spans, a pretrained Matryoshka sentence \textbf{embedder} (sharpened by a light contrastive fine-tuning and an error-driven hard-negative mining pass) maps each span to a vector, and a \textbf{typing head} assigns a type.
The pipeline is modular: mention detector, embedder and typing head are independently swappable. We vary each in the ablations, testing many combinations to reach the best pipeline: closed-set detectors (BERT \cite{devlin2019bert}, RoBERTa \cite{liu2019roberta}) against GLiNER \cite{zaratiana2024gliner} at three sizes, several embedders (BGE \cite{xiao2024cpack}, MPNet \cite{song2020mpnet}, mxbai \cite{li2024aoe}, E5-base \cite{wang2022e5} and Nomic \cite{nussbaum2024nomic}), the best Matryoshka \cite{kusupati2022matryoshka} dimension to trade quality for compute, and different algorithms for entity typing.

The typing head exposes two operating points: \textbf{OPENER-ZS}, the label-free zero-shot head we put forward as the open-world default, and \textbf{OPENER-Sup}, a supervised upper bound for domains whose entity classes are fixed and labelled. We detail both below.

Our contributions are as follows.
\begin{itemize}
\item \textbf{The OPENER pipeline, a new modular architecture}: to the best of our knowledge, it is the first open-world NER pipeline assembled \emph{entirely} from ready-to-use, pretrained components, with no encoder trained from scratch. Instead, it relies on \emph{transfer learning} through a frozen mention detector and a Matryoshka embedder, where only the latter undergoes light fine-tuning. The pipeline's three stages (mention detection, embedding, and typing) are \emph{independently swappable}, and the typing head itself supports two interchangeable modes (zero-shot or supervised). We consider this plug-and-play modularity to be a novel contribution to open-world NER.
The embedder is lightly refined in two steps: an initial contrastive fine-tuning to cluster same-type mentions, followed by an \emph{error-driven hard-negative mining} pass targeting the most frequently confused type pairs. This second step improves typing-on-gold AMI by $8.5$ points compared to standard contrastive fine-tuning, successfully demonstrating that a general-purpose pretrained representation transfers well enough to type open-world entities.
\item \textbf{OPENER-ZS}: it is a label-free typing head that needs no target labels and combines three signals. Here is the method:

(I) \emph{Label-name prototypes}: we embed each candidate type \emph{name} (the words ``person'', ``location'', \dots) into a prototype vector and assign a mention to its nearest prototype: so ``Marie Curie'' is typed \textsc{person} because her mention vector lands closest to the embedding of the word ``person''.

(II) A \emph{transductive refinement}: the prototypes are re-estimated from the test mentions themselves (not just the label names), tightening each type to the domain at hand without using any labels.

(III) A \emph{fusion} with the detector's own zero-shot type guess, GLiNER already proposes a type when it detects ``Marie Curie'', a signal that prototype matching alone would discard. It is the best zero-shot system end-to-end on our benchmark, ahead of GLiNER-L, GNER and OWNER under an identical protocol.
\item \textbf{OPENER-Sup}: a supervised operating point for domains whose entity classes are fixed and labelled. Where OPENER-ZS matches each mention to a label \emph{name}, OPENER-Sup instead \emph{learns} the geometry of each type from labelled examples. Here is the idea: once mentions are embedded, a linear classifier carves the high-dimensional embedding space into regions, one per entity type, separated by hyperplanes. A new span is then typed by which region its vector falls into: a mention landing in the ``chemical'' region is typed \textsc{chemical}, because same-type mentions cluster together in that space. We use a \emph{balanced} linear support-vector classifier (LinearSVC), where ``balanced'' re-weights each class inversely to its frequency so that rare types are not drowned out by frequent ones, a common issue since entity types are heavily imbalanced. This trades openness for accuracy: it is the most accurate configuration in our benchmark, the best end-to-end AMI of all compared systems and the best typing on gold mentions, but it needs target labels.
\item \textbf{An latency- and energy-aware benchmark} of open-world NER on $13$ datasets that reports quality, latency and energy jointly, separates typing-on-gold from end-to-end, and shows that the dominant open-world bottleneck is mention detection rather than entity typing. This is why an \emph{open-world} detector is essential: closed-set NER detectors (BERT/RoBERTa) collapse on specialized domains, an efficiency/quality gap an accuracy-only evaluation would hide. The comparison also includes a small instruction-tuned LLM, Qwen2.5-1.5B-Instruct in 4-bit (the largest generative model that fits our commodity GPU), which turns out to be both the least accurate and the most expensive of all systems (on a $150$-sentence subsample, \autoref{tab:main}). All OPENER results are averaged over three full embedder-retraining seeds ($7$, $42$, $123$) and reported as mean\,$\pm$\,std. The compared baselines are frozen checkpoints whose inference is deterministic. The differences we report are therefore robust to initialisation.
\end{itemize}

The rest of the paper is organised as follows.
\autoref{sec:related} reviews related work, \autoref{sec:method} details the OPENER pipeline, \autoref{sec:experiments} presents the benchmark and ablations, and \autoref{sec:conclusion} concludes.
