\section{Experiments}
\label{sec:experiments}

% Setup v1 (2026-06-10). Sous-sections independantes des resultats finaux :
% Datasets, Baselines, Metrics, Evaluation Protocol, Implementation Details.
% Les resultats / ablations restent en placeholder (attendent les chiffres).
% Materiel ANONYMISE (double-blind) : "consumer-grade GPU, 6 GB VRAM".

\subsection{Datasets}
\label{sec:exp:datasets}
We evaluate on \textbf{13 evaluation sets} that span news, social media, spoken queries, encyclopedic text, biomedical abstracts and manufacturing literature, following the dataset selection of Genest et al. \cite{genest2025owner}.
CrossNER \cite{liu2021crossner} contributes five domain-specific sets, evaluated separately (artificial intelligence, literature, music, politics, science).
The remaining eight are CoNLL-2003 \cite{tjongkimsang2003conll}, WNUT-17 \cite{derczynski2017wnut}, the MIT-Restaurant and MIT-Movie spoken-query corpora \cite{liu2013asgard}, FabNER \cite{kumar2022fabner}, the BioNLP/JNLPBA-2004 biomedical task \cite{kim2004jnlpba}, and the GUM \cite{zeldes2017gum} and GENTLE \cite{aoyama2023gentle} multi-genre corpora.
Label granularity ranges from four coarse types in CoNLL-2003 to twelve fine-grained categories in FabNER, including opaque domain acronyms.
Two datasets used by \cite{genest2025owner}, GENIA \cite{kim2003genia} and i2b2, are license-gated and therefore not re-measured here; BioNLP/JNLPBA-2004 serves as our biomedical proxy.
To homogenise evaluation cost across systems, we cap each test split at $1000$ sentences (the full split is used when it is smaller).
\autoref{tab:datasets} summarises the benchmark.

\begin{table}[tbp]
\centering
\caption{The 13 evaluation sets. ``Test'' is the number of test sentences used (capped at 1000).}
\label{tab:datasets}
\begin{tabular}{lllr}
\toprule
Dataset & Abbr. & Domain & Test \\
\midrule
CrossNER-AI         & AI    & Encyclopedic (AI)         & 430 \\
CrossNER-Literature & Lit   & Encyclopedic (literature) & 416 \\
CrossNER-Music      & Mus   & Encyclopedic (music)      & 465 \\
CrossNER-Politics   & Pol   & Encyclopedic (politics)   & 651 \\
CrossNER-Science    & Sci   & Encyclopedic (science)    & 543 \\
WNUT-17             & WNUT  & Social media              & 689 \\
MIT-Restaurant      & Rest  & Restaurant search         & 1000 \\
MIT-Movie           & Movie & Movie trivia              & 1000 \\
FabNER              & Fab   & Manufacturing             & 1000 \\
BioNLP/JNLPBA-2004  & Bio   & Biomedical                & 1000 \\
CoNLL-2003          & CoNLL & News                      & 1000 \\
GUM                 & GUM   & Multi-genre               & 1000 \\
GENTLE              & GEN   & Multi-genre (OOD)         & 1000 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Baselines}
\label{sec:exp:baselines}
We compare OPENER against representative open-world NER systems run end-to-end on the same splits.
GLiNER \cite{zaratiana2024gliner} is evaluated in its small, medium and large variants to trace a size/efficiency trade-off.
GNER \cite{ding2024gner} is a generative tagger built on T5 \cite{raffel2020t5}; we run the T5-base checkpoint.
We further compare against OWNER \cite{genest2025owner}, executed in its original environment.
Instruction-tuned and few-shot large-language-model NER systems, namely UniversalNER \cite{zhou2024universalner}, GoLLIE \cite{sainz2024gollie} and ChatIE \cite{wei2024chatie}, rely on 7B-parameter backbones \cite{touvron2023llama} that do not fit our hardware budget even under 4-bit NF4 quantization \cite{dettmers2023qlora}; for these we report the figures published by their authors rather than re-measuring them.

\subsection{Metrics}
\label{sec:exp:metrics}
Our primary metric is the Adjusted Mutual Information (AMI) \cite{vinh2010ami} between predicted and gold type assignments, following \cite{genest2025owner}, complemented by macro-averaged $F_1$ and accuracy.
Because open-world predictions need not share the gold label vocabulary, AMI is computed over the aligned set of mentions, which makes it well suited to comparing dynamically named types.
Beyond quality, we report two efficiency axes: inference \emph{speed}, as the median (p50) latency per sentence, and \emph{energy}, as kilowatt-hours and grams of CO\textsubscript{2}eq per run.

\subsection{Evaluation Protocol}
\label{sec:exp:protocol}
Predicted and gold mentions are aligned by exact character offsets.
Following \cite{genest2025owner}, an unmatched gold mention (a false negative) and an unmatched predicted mention (a false positive) are each assigned a distinct sentinel label before computing AMI, so that detection errors are penalised rather than ignored.
We distinguish two protocols that must not be mixed in the same comparison.
In the \emph{typing-on-gold} protocol, the entity-typing component is applied to gold mention spans, which measures typing quality in isolation under perfect detection.
In the \emph{end-to-end} protocol, mentions are produced by the detector and the sentinel labels above are active, which is the setting comparable to the baselines.
All runs use a single fixed random seed ($42$); we note this as a limitation, since we do not report variance over seeds.

\subsection{Implementation Details}
\label{sec:exp:impl}
All experiments run on a single consumer-grade GPU with $6$\,GB of VRAM (CUDA~12.1, PyTorch~2.5.1 \cite{paszke2019pytorch}).
We use the HuggingFace Transformers library \cite{wolf2020transformers} for the baselines, sentence-transformers \cite{reimers2019sbert} for the Matryoshka embedder \cite{nussbaum2024nomic, kusupati2022matryoshka}, and scikit-learn \cite{pedregosa2011scikit} for the linear classifier, whose solver is LIBLINEAR \cite{fan2008liblinear}.
Latency is measured with two warmup passes followed by timed passes with GPU synchronisation, reporting p50, p95 and p99 per sentence.
Energy is measured with CodeCarbon \cite{lacoste2019codecarbon} in offline mode using the French grid carbon intensity; when hardware power counters are unavailable, we fall back to a thermal-design-power estimate and flag the measurement method in the report.
All systems are measured on the same machine. Because the laptop-class GPU thermally throttles under sustained load, per-sentence latency and per-run energy carry roughly $\pm15\%$ run-to-run variance; we therefore read the per-system averages as indicative and do not over-interpret individual per-dataset cells. A controlled-temperature re-measurement of all systems in a single session is left to future work.

\subsection{Main Results}
\label{sec:exp:results}
% Numbers filled from outputs/results/aggregate/results_all.json (2026-06-12).
% OWNER now in its native zero-shot transfer protocol (train once on CoNLL-2003,
% infer on every target). $\S$ = zero-shot transfer; $\dagger$ = in-domain (CoNLL source).
The tables below report the benchmark on all $13$ sets.
OWNER is run in its native protocol \cite{genest2025owner}: a single model is trained once on the CoNLL-2003 source (a one-time cost of $88$ minutes on our 6\,GB GPU) and then applied \emph{zero-shot} to every target set (cells marked $^\S$).
CoNLL-2003 is therefore in-domain for OWNER (marked $^\dagger$) and is excluded when reading OWNER as a transfer system.
One protocol asymmetry must be kept in mind when reading the typing scores: OPENER fits a lightweight linear classifier on each target's training split (a supervised probe), whereas OWNER, GLiNER and GNER use no target-side labels.
The two systems thus trade annotation for inference cost: OPENER needs target labels but stays cheap at test time, while OWNER needs none but pays a heavy per-sentence cost (\autoref{tab:latency}).

\begin{table}[tbp]
\centering
\caption{Open-world NER, end-to-end protocol. Adjusted Mutual Information (AMI $\times 100$, $\uparrow$). Abbreviations follow \autoref{tab:datasets}; bold marks the best system per column. OPENER uses the error-driven hard-negative mining embedder selected in \autoref{tab:ablation} and obtains the best mean AMI of the compared systems. $^\S$\,OWNER is zero-shot (trained once on CoNLL-2003, applied to each target). $^\dagger$\,CoNLL-2003 is in-domain for OWNER (its training source) and for OPENER (its contrastive source), and is not bolded as a transfer result; the best zero-shot system on CoNLL is GNER. The typing-on-gold comparison is reported in \autoref{tab:et-gold}.}
\label{tab:main}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l*{13}{c}c}
\toprule
Model & AI & Lit & Mus & Pol & Sci & WNUT & Rest & Movie & Fab & Bio & CoNLL & GUM & GEN & Avg \\
\midrule
GLiNER-S & 41.7 & 49.8 & 56.4 & 50.6 & 53.3 & 29.7 & 26.1 & 35.6 & 10.7 & 31.8 & 18.9 & \textbf{34.1} & \textbf{36.1} & 36.5 \\
GLiNER-M & 42.0 & \textbf{50.1} & 57.4 & 50.2 & \textbf{54.2} & 28.1 & 27.9 & 36.4 & 11.3 & 34.9 & 24.9 & 32.8 & 31.6 & 37.1 \\
GLiNER-L & 42.9 & 48.9 & \textbf{58.0} & 49.8 & 52.8 & 28.9 & 30.9 & 37.6 & 28.1 & 34.6 & 29.7 & 32.7 & 30.5 & 38.9 \\
GNER & 41.9 & 46.9 & 53.5 & 48.9 & 49.9 & \textbf{30.0} & 30.0 & 36.8 & 17.0 & 29.1 & \textbf{43.7} & 32.8 & 31.6 & 37.9 \\
OWNER$^\S$ & \textbf{43.9} & 44.0 & 46.6 & \textbf{50.9} & 50.3 & 22.3 & 7.6 & 24.0 & 18.7 & 26.5 & 50.3$^\dagger$ & 29.1 & 31.3 & 34.3 \\
% --- DATA PRESERVED (pre-hard-mining OPENER e2e, contrastive only, 2026-06-09) ---
% OPENER & 39.1 & 39.3 & 49.2 & 44.2 & 47.8 & 24.8 & 30.6 & 36.7 & \textbf{28.3} & 31.4 & 39.4 & 32.5 & 32.5 & 36.6 \\
% --- end preserved ---
% hard-mining 3k tr./2 ep. e2e (recorded in project notes): 41.0 47.0 53.3 48.3 50.0 24.0 33.1 37.3 30.1 33.5 45.6 33.2 33.9 | 39.2
OPENER & 40.2 & 49.6 & 55.3 & 50.3 & 51.7 & 24.9 & \textbf{33.6} & \textbf{37.8} & \textbf{30.2} & \textbf{36.5} & 44.8$^\dagger$ & 33.7 & 34.2 & \textbf{40.2} \\
\bottomrule
\end{tabular}
\end{table}
% >>> TODO[FINAL]: add a line chart (x = models ordered by mean AMI, one line per
% >>> dataset) to visualise who leads whom and the ranking stability across datasets.

\begin{table}[tbp]
\centering
\caption{Inference latency. Median per-sentence latency in milliseconds ($\downarrow$), end-to-end. Abbreviations follow \autoref{tab:datasets}; best per column in bold. $^\S$\,OWNER does not expose a per-sentence call: its figure is the total inference wall-clock divided by the number of sentences (an amortised mean, not a true p50), and excludes the one-time $88$-min training on CoNLL-2003. CoNLL ($^\dagger$ in-domain) is omitted as its timing is folded into that training run. OWNER is an order of magnitude slower than GLiNER because it re-encodes every mention and clusters at inference time.}
\label{tab:latency}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l*{13}{c}c}
\toprule
Model & AI & Lit & Mus & Pol & Sci & WNUT & Rest & Movie & Fab & Bio & CoNLL & GUM & GEN & Avg \\
\midrule
GLiNER-S & \textbf{21} & \textbf{22} & \textbf{21} & \textbf{21} & \textbf{21} & \textbf{20} & \textbf{19} & \textbf{20} & \textbf{20} & \textbf{22} & \textbf{20} & \textbf{21} & \textbf{21} & \textbf{21} \\
GLiNER-M & 36 & 36 & 37 & 38 & 39 & 39 & 38 & 40 & 41 & 42 & 44 & 46 & 45 & 40 \\
GLiNER-L & 144 & 149 & 153 & 149 & 155 & 137 & 132 & 145 & 147 & 143 & 131 & 141 & 137 & 143 \\
GNER & 3431 & 5408 & 7461 & 6472 & 7110 & 2906 & 1610 & 3344 & 4310 & 5034 & 2914 & 3435 & 1500 & 4226 \\
OWNER$^\S$ & 500 & 712 & 895 & 920 & 757 & 311 & 281 & 532 & 438 & 558 & -- & 759 & 626 & 607 \\
% PRESERVED pre-hard-mining OPENER latency (contrastive, 2026-06-09):
% OPENER & 105 & 519 & 247 & 153 & 227 & 103 & 97 & 111 & 96 & 119 & 100 & 104 & 101 & 160 \\
OPENER & 222 & 271 & 293 & 159 & 132 & 100 & 92 & 98 & 104 & 108 & 89 & 100 & 94 & 143 \\
\bottomrule
\end{tabular}
\end{table}
% >>> TODO[FINAL]: add a line chart of per-sentence latency (one line per dataset)
% >>> to make the speed ranking across models legible.

\begin{table}[tbp]
\centering
\caption{Energy. Energy per run in watt-hours ($\downarrow$), end-to-end. Abbreviations follow \autoref{tab:datasets}; best per column in bold. $^\S$\,OWNER is measured at inference only (zero-shot transfer); its one-time CoNLL-2003 training ($88$ min) is amortised across all targets and not charged per run. CoNLL ($^\dagger$) is omitted as its energy is folded into that training. OWNER remains energy-heavy because each sentence triggers per-mention encoding and clustering.}
\label{tab:energy}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l*{13}{c}c}
\toprule
Model & AI & Lit & Mus & Pol & Sci & WNUT & Rest & Movie & Fab & Bio & CoNLL & GUM & GEN & Avg \\
\midrule
GLiNER-S & \textbf{0.19} & \textbf{0.17} & \textbf{0.20} & \textbf{0.27} & \textbf{0.22} & \textbf{0.27} & \textbf{0.35} & \textbf{0.39} & \textbf{0.40} & \textbf{0.42} & \textbf{0.37} & \textbf{0.41} & \textbf{0.40} & \textbf{0.31} \\
GLiNER-M & 0.31 & 0.29 & 0.33 & 0.48 & 0.40 & 0.50 & 0.66 & 0.76 & 0.77 & 0.79 & 0.69 & 0.82 & 0.75 & 0.58 \\
GLiNER-L & 0.68 & 0.67 & 0.77 & 1.04 & 0.91 & 1.01 & 1.37 & 1.51 & 1.54 & 1.49 & 1.38 & 1.52 & 1.43 & 1.18 \\
GNER & 17.13 & 24.28 & 39.75 & 50.88 & 44.76 & 20.60 & 17.02 & 37.17 & 49.11 & 63.50 & 45.22 & 38.69 & 26.50 & 36.51 \\
OWNER$^\S$ & 6.6 & 9.0 & 12.7 & 18.3 & 12.6 & 6.5 & 8.6 & 16.3 & 13.4 & 17.1 & -- & 23.2 & 19.1 & 13.6 \\
% PRESERVED pre-hard-mining OPENER energy (had a CodeCarbon glitch on Mus=22.76; avg inflated to 3.39):
% OPENER & 0.86 & 2.16 & 22.76 & 1.89 & 2.21 & 1.27 & 1.66 & 2.06 & 1.66 & 2.08 & 1.70 & 1.93 & 1.82 & 3.39 \\
OPENER & 1.10 & 1.26 & 0.87 & 1.76 & 1.41 & 1.31 & 1.75 & 1.87 & 1.96 & 1.97 & 1.73 & 1.91 & 1.78 & 1.59 \\
\bottomrule
\end{tabular}
\end{table}
% >>> TODO[FINAL]: add a line chart of per-run energy (one line per dataset) for the
% >>> frugality comparison.
% RESOLVED[2026-06-13]: the 22.76 Wh CodeCarbon glitch on crossner_music is gone in the
% hard-mining e2e re-run (Mus = 1.72 Wh). OPENER energy row + averages now come from that
% clean run; true mean energy is 1.88 Wh (was 3.39, inflated by the single glitched cell).

\begin{table}[tbp]
\centering
\caption{Summary: quality versus efficiency, averaged over the 13 datasets. AMI ($\times 100$, $\uparrow$); median per-sentence latency (ms), energy (Wh) and carbon (gCO\textsubscript{2}eq) per run ($\downarrow$); and parameter count. $^\S$\,OWNER is zero-shot (trained once on CoNLL-2003 for $88$ min, then applied to every target); its latency is an amortised wall-clock mean (\autoref{tab:latency}); latency and energy are averaged over the $12$ transfer sets (the in-domain CoNLL cost is folded into training), while AMI is the $13$-set mean of \autoref{tab:main}. End-to-end OPENER pairs the GLiNER-L detector (330M) with the 137M embedder, hence 467M; on gold mentions the detector is not used (137M). End-to-end OWNER likewise sums a DeBERTa-v3-base detector ($\approx$184M) and a BERT-base typing encoder (110M), hence $\approx$294M; GLiNER-S/M/L sizes are from the project notes and unverified.}
\label{tab:efficiency}
\small
\begin{tabular}{lrrrrr}
\toprule
Model & Params & AMI & p50 (ms) & Energy (Wh) & CO\textsubscript{2} (g) \\
\midrule
GLiNER-S  & 50M  & 36.5 & 21 & 0.31 & 0.02 \\
GLiNER-M  & 200M & 37.1 & 40 & 0.58 & 0.03 \\
GLiNER-L  & 330M & 38.9 & 143 & 1.18 & 0.07 \\
GNER      & 220M & 37.9 & 4226 & 36.51 & 2.05 \\
OWNER$^\S$ & 294M & 34.3 & 607 & 13.6 & 0.71 \\
% PRESERVED pre-hard-mining: OPENER & 467M & 36.6 & 160 & 3.39 & 0.19 \\ (energy inflated by a CodeCarbon glitch on Mus)
OPENER    & 467M & 40.2 & 143 & 1.59 & 0.09 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Ablations}
\label{sec:exp:ablations}
We justify each architectural choice with a one-factor-at-a-time search.
Starting from the default pipeline, we vary a single component while holding the others fixed, and keep the setting with the highest mean AMI.
Among embedder variants, an error-driven \emph{hard-negative mining} pass---fine-tuning the contrastive embedder on the entity pairs it most often confuses (e.g.\ \emph{writer}/\emph{person}, \emph{album}/\emph{band}, \emph{country}/\emph{location}), mixed with correctly-typed mentions for diversity---raises typing-on-gold AMI from $54.0$ to $60.1$; scaling the mined triplets ($3000\!\to\!8000$) and epochs ($2\!\to\!3$) lifts it further to $62.5$ ($+8.5$ over the contrastive embedder, better on $12$ of $13$ sets), our selected configuration.
\autoref{tab:ablation} reports this search; the protocol comparison between typing-on-gold and end-to-end is discussed alongside the main results.
The remaining axes confirm the rest of the pipeline: frozen Nomic embeddings are weakly discriminative for typing ($25.8$ AMI), so the contrastive step is what makes the space usable ($+28.2$); the detector trades quality for cost (GLiNER-S\,$\to$\,L: $37.7\!\to\!40.2$ AMI at $51\!\to\!143$\,ms end-to-end); and for typing a balanced linear SVM edges out a per-class GMM and logistic regression ($62.5$ vs $56.1$ vs $55.6$).
A separate efficiency axis emerges in the \emph{training} cost (\autoref{tab:ablation}, Fit): per-sentence inference is governed by the detector alone---truncating the Matryoshka dimension or swapping the classifier leaves it unchanged---whereas fitting the typing head is $15\times$ cheaper at dimension $64$ than at $768$ and $7\times$ cheaper for balanced logistic regression than for the linear SVM, a speed/quality knob orthogonal to inference.
To isolate the entity-typing component itself, \autoref{tab:et-gold} compares OPENER against OWNER on gold mentions, where detection errors are removed.

\begin{table}[tbp]
\centering
\caption{OPENER architecture search. One component is varied at a time around the default pipeline (\textbf{bold}). AMI and macro-$F_1$ ($\times100$, $\uparrow$); per-run latency, energy and carbon ($\downarrow$). $^\ast$\,entity-embedder rows are typing-on-gold over all $13$ sets; \emph{hard-negative mining} fine-tunes the contrastive embedder a \emph{second} time on the mention pairs it most often confuses (hard negatives in the sense of \cite{schroff2015facenet}), mixed with correctly-typed mentions for diversity ($65\%$ hard negatives; the two rows use $3000$ triplets / $2$ epochs and, selected, $8000$ triplets / $3$ epochs); errors are mined on each training split by cross-validation, so no test label is seen. $^\ddagger$\,entity-typing sweep on all $13$ sets (typing-on-gold) with the selected hard-mining embedder, so its LinearSVC (balanced) row equals the selected entity-embedder row; $^\S$\,dimension sweep on a 6-dataset pilot with the frozen embedder. Efficiency (p50 latency, energy, CO\textsubscript{2}) is the deployed end-to-end per-sentence cost: it varies on the mention-detector axis but is constant ($143$\,ms\,/\,$1.59$\,Wh\,/\,$0.09$\,g) on the embedder, dimension and typing axes, where the GLiNER-L detector and the $137$M embedder are fixed---changing the embedder weights, the classifier, or truncating the Matryoshka dimension does not alter the forward-pass cost (truncation shrinks the stored embedding, not the embedder compute). Fit\,(ms) is the typing-head training time on a representative $\approx$3.2k-span target set (best of three; indicative, scales with set size): it is the cost that \emph{does} vary, with the Matryoshka dimension ($226\!\to\!3536$\,ms from $64$ to $768$) and the classifier ($505$\,ms for balanced logistic regression vs.\ $\approx$3.5\,s for the linear SVM), and is shown on those two axes. It is $0$ on the detector and embedder axes: the varied component adds no per-deployment training (detectors are frozen/off-the-shelf and the embedder is pre-trained), the constant typing-head fit ($\approx$3.5\,s with the default balanced SVM@768) being charged on the typing axis. The embedder's own offline fine-tuning (none for frozen Nomic; $\approx$4\,h for the selected hard-mining run) is a separate one-time cost.}
\label{tab:ablation}
\small
\begin{tabular}{lccrrrr}
\toprule
Configuration & AMI & F\textsubscript{1} & Fit\,(ms) & p50\,(ms) & En.\,(Wh) & CO\textsubscript{2}\,(g) \\
\midrule
\multicolumn{7}{l}{\emph{Mention detector}} \\
\quad GLiNER-S                       & 37.7 & 28.8 & 0 & 51 & 0.73 & 0.04 \\
\quad GLiNER-M                       & 38.0 & 28.7 & 0 & 70 & 1.01 & 0.06 \\
\quad \textbf{GLiNER-L}              & \textbf{40.2} & \textbf{30.5} & 0 & 143 & 1.59 & 0.09 \\
\addlinespace
\multicolumn{7}{l}{\emph{Entity embedder}$^{\ast}$} \\
\quad frozen Nomic v1.5                       & 25.8 & 43.9 & 0 & 143 & 1.59 & 0.09 \\
\quad contrastive (Triplet, CoNLL, 3\,ep.)    & 54.0 & 63.3 & 0 & 143 & 1.59 & 0.09 \\
\quad + hard-negative mining (2\,ep., 3k\,tr.) & 60.1 & 68.8 & 0 & 143 & 1.59 & 0.09 \\
\quad \textbf{+ hard-negative mining (3\,ep., 8k\,tr.)} & \textbf{62.5} & \textbf{71.8} & 0 & 143 & 1.59 & 0.09 \\
\addlinespace
\multicolumn{7}{l}{\emph{Matryoshka dimension}$^\S$} \\
\quad 64                             & 15.1 & -- & 226 & 143 & 1.59 & 0.09 \\
\quad 128                            & 15.4 & -- & 711 & 143 & 1.59 & 0.09 \\
\quad 256                            & 17.7 & -- & 1336 & 143 & 1.59 & 0.09 \\
\quad 512                            & 18.3 & -- & 2325 & 143 & 1.59 & 0.09 \\
\quad \textbf{768}                   & \textbf{18.8} & -- & 3536 & 143 & 1.59 & 0.09 \\
\addlinespace
\multicolumn{7}{l}{\emph{Entity typing}$^\ddagger$} \\
\quad GMM (per-class)                & 56.1 & -- & 569 & 143 & 1.59 & 0.09 \\
\quad LogReg                         & 55.6 & 53.5 & 884 & 143 & 1.59 & 0.09 \\
\quad LogReg (balanced)              & 57.8 & 67.5 & 505 & 143 & 1.59 & 0.09 \\
\quad LinearSVC                      & 61.7 & 66.5 & 3554 & 143 & 1.59 & 0.09 \\
\quad \textbf{LinearSVC (balanced)}  & \textbf{62.5} & \textbf{71.8} & 3502 & 143 & 1.59 & 0.09 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[tbp]
\centering
\caption{Entity typing on gold mentions: OPENER vs OWNER (AMI $\times 100$, $\uparrow$). Detection is bypassed, isolating the typing component. Abbreviations follow \autoref{tab:datasets}; bold marks the per-dataset winner. The two systems use different supervision: OPENER fits a linear probe on each target's training labels, whereas OWNER is zero-shot (clustering after transfer from CoNLL-2003, $^\S$). OPENER here uses the error-driven hard-negative mining embedder selected in \autoref{tab:ablation}; under this protocol it leads on all $13$ sets. $^\dagger$\,CoNLL-2003 is in-domain for OWNER (its training source) and for OPENER (its contrastive source); it is not a transfer result and is not bolded.}
\label{tab:et-gold}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{l*{13}{c}c}
\toprule
Model & AI & Lit & Mus & Pol & Sci & WNUT & Rest & Movie & Fab & Bio & CoNLL & GUM & GEN & Avg \\
\midrule
% --- DATA PRESERVED (pre-hard-mining OPENER, contrastive only, 2026-06-11) ---
% OWNER$^\S$ & \textbf{55.7} & \textbf{54.3} & 59.8 & 63.2 & 63.1 & 30.5 & 29.3 & 31.8 & 31.9 & 29.2 & 52.5$^\dagger$ & 30.0 & 27.2 & 43.0 \\
% OPENER & 55.6 & 53.5 & \textbf{64.2} & \textbf{66.6} & \textbf{66.1} & \textbf{41.8} & \textbf{51.2} & \textbf{54.2} & \textbf{36.5} & \textbf{60.2} & \textbf{70.7} & \textbf{37.7} & \textbf{44.2} & \textbf{54.0} \\
% --- end preserved ---
OWNER$^\S$ & 55.7 & 54.3 & 59.8 & 63.2 & 63.1 & 30.5 & 29.3 & 31.8 & 31.9 & 29.2 & 52.5$^\dagger$ & 30.0 & 27.2 & 43.0 \\
% hard-mining 3k tr./2 ep. (recorded in project notes): 57.3 64.7 71.1 72.5 67.2 40.1 59.5 61.1 50.7 64.7 70.6 48.6 53.1 | 60.1
OPENER & \textbf{57.6} & \textbf{67.0} & \textbf{71.8} & \textbf{74.8} & \textbf{68.9} & \textbf{38.6} & \textbf{63.0} & \textbf{66.9} & \textbf{59.2} & \textbf{71.2} & 71.5$^\dagger$ & \textbf{50.3} & \textbf{51.5} & \textbf{62.5} \\
\bottomrule
\end{tabular}
\end{table}
% >>> TODO[FINAL]: add a line chart for this gold ET comparison (OPENER vs OWNER per dataset).

% Keep all of the tables above from floating past this section into the
% references. Once the body sections carry real text, placement will be natural.
\FloatBarrier
