\section{Conclusion}
\label{sec:conclusion}

We presented OPENER, an open-world NER pipeline built from ready-to-use components  --  a frozen detector, a pretrained Matryoshka embedder sharpened by a light contrastive fine-tuning with error-driven hard-negative mining, and a swappable typing head  --  with no encoder trained from scratch.
On gold mentions the contrastive space is highly discriminative, and a balanced linear probe, OPENER-Sup, is the most accurate system in our benchmark, leading on all $13$ sets and surpassing a zero-shot transfer of OWNER ($62.5$ vs $43.0$ AMI); it also attains the best end-to-end AMI of the compared systems ($40.2$), at the cost of target labels.
Removing that supervision, our zero-shot head OPENER-ZS  --  label-name prototypes refined transductively and fused with the detector's own predictions  --  is the best of the compared zero-shot systems end-to-end ($39.4$), ahead of GLiNER-L, GNER and OWNER, at one to two orders of magnitude lower energy than the trained-encoder and generative baselines.

Two findings carry beyond the specific system.
First, an open-world typing head need not train its own encoder: a general-purpose embedding, once sharpened by a light contrastive step, is already discriminative enough to type entities, and the cost that does matter is fitting the cheap typing head, not the embedder.
Second, the three-axis evaluation locates the open-world bottleneck in \emph{detection} rather than typing: typing on gold mentions reaches $62.5$ AMI but end-to-end scores drop to $40.2$ as detector recall falls on specialised domains, so improving the detector -- not the typing head -- is the clearest axis for future gains. The evaluation also exposes efficiency gaps that an accuracy-only comparison would hide.

The approach has clear limits.
End-to-end quality is capped by detector recall on specialised domains, the shared bottleneck, not by the typing head.
Supervised typing needs target-domain labels, and while OPENER-ZS removes them at typing time it depends on the quality of the label names and, in its transductive form, on access to the unlabelled test mentions; the inductive zero-shot head is correspondingly weaker.
Finally, our experiments run on deliberately modest hardware: a single consumer GPU with $6$\,GB of VRAM, subject to thermal throttling.
This is the very setting that makes OPENER deployable, but it also bounded our experimental budget: despite an already broad benchmark ($13$ datasets, three axes and extensive ablations), we train the embedder with a single random seed and cap each test split at $1000$ sentences. Because inference is deterministic, the quality scores themselves carry no run-to-run noise, and a test-set bootstrap confirms they are stable (95\% CIs of $\pm 1$--$1.5$ AMI, \autoref{tab:bootstrap}); only latency and energy fluctuate ($\pm 15\%$) and are therefore read as averages rather than precise points.
A larger compute budget would extend the same protocol to multiple embedder-\emph{retraining} seeds with significance tests, the full uncapped test splits, and a controlled-temperature re-measurement of latency and energy.

Future work includes improving mention detection on specialised domains to lift the end-to-end ceiling, an inductive zero-shot head that does not rely on the test distribution, a visualisation of the contrastive type geometry (e.g.\ UMAP or t-SNE), and extending the decoupled detect-then-type design beyond NER to entity linking and relation extraction.
